Pool Model

### 2025-08-26 pool_model.tex
rename bib file
Supplement1: math formalism, do some rewritings

\usepackage[utf8]{inputenc}
\usepackage[a4paper,
            inner=2.5cm,
            outer=3cm
            ]{geometry}
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off, font=small]{caption}            
\usepackage{acronym}
\usepackage{graphicx}% Include figure files
\usepackage{bm}% bold math
\usepackage{color}
\usepackage{hyperref}
\usepackage[backend=biber,
            sorting=none,
%            citestyle=numeric-comp,
%\            bibstyle=ieee, 
            style=numeric-comp,
            hyperref=true,
            maxnames=2,
            url=false,
            isbn=false,
            eprint=false,
            date=year,
            doi=false,
            giveninits=true
            ]{biblatex}
\addbibresource{Bacteria Modelling.bib}
\usepackage{xpatch}
\xpatchbibmacro{journal+issuetitle}{%
\usebibmacro{journal}%
\setunit*{\addspace}%
}{%
\usebibmacro{journal}%
\setunit*{\addcomma\space}%
}{}{}

@@

% ================================================================================================
\beginsupplement
\section{Supplementary Material}
% ================================================================================================
\subsection{General Mathematical Formalism}  \label{ssec:supplement1}
Consider the system consisting of $N$ species in $M$ different states.
In the general representation, the system can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
\begin{equation}
   \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
\label{eq:model_ODE_x}
\end{equation}
where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector describing bacterial count.
In matrix form it can be written as
\begin{equation}
    \mbx = \begin{pmatrix}
        x^1_1  & \dots & x^1_N  \\
        \vdots &       & \vdots \\
        x^M_1  & \dots & x^M_N  \\
            \end{pmatrix}
    \label{eq:model_bact}
\end{equation}
with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
\begin{equation}
    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu).
    \label{eq:model_microenv}
\end{equation}
Here function $\mbH$ is in general case also non-linear.\\
%
The advantage of the following formulation is its generality.
Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.\\
%
To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
\begin{equation}
    N_i = \sum_k \mbx_i^k.
\label{eq:N_pro_species}
\end{equation}
While the total bacterial count of all species of our system is
\begin{equation}
    N = \sum_i \sum_k \mbx_i^k.
\label{eq:N_total}
\end{equation}
%
Using such matrix representation of the bacterial system, we can rewrite introduced above system in more compact form.
Here the concentrations of the pools are components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
The three pool system described with equations~\ref{eq:ode_3pools} will be written as
\begin{equation}
    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
        -(\mu + \lambda) & 0             & 0      \\
        \lambda          & \alpha - \mu' & 0      \\
        \mu              & \mu'          & -\beta 
    \end{pmatrix}.
\label{eq:x_repres_3pools}
\end{equation}
%
With introduction of the limiting resource the system dynamics presented by the \acp{ode}~\ref{eq:ode_3pool_resource} will be rewritten in the following non-linear system:
\begin{eqnarray}
    \dot \mbx  &=& \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
        -(\mu + \lambda) & 0               & 0      \\
        \lambda          & \alpha R - \mu' & 0      \\
        \mu              & \mu'            & -\beta 
    \end{pmatrix} \mbx\\
    \dot \mbm &=& \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
\label{eq:x_repres_3pools_resource}
\end{eqnarray}
\begin{equation}
    \dot \mbx = \begin{pmatrix}
        -(\mu + \lambda) & 0       & 0 \\
         \lambda         & \alpha  & 0 \\
         \mu &  \mu'& - \beta 
    \end{pmatrix} 
    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
\label{eq:x_repres_3pools_resource2}
\end{equation}
Here the first term is the linear part and the second term define the quadratic behavior of the system.
%
\paragraph{Generalized Lotka-Volterra}

This system of equations defines the microenvironment $\mbm$ and bacteria state matrix $\mbx$ as
\begin{equation}
    \mbm = R, \hspace{1cm}
    \mbx = \begin{pmatrix}
        L_A & L_B \\
        G_A & G_B 
    \end{pmatrix}.
\label{eq:gLV_x_m}
\end{equation}
%

>>>>

\usepackage[utf8]{inputenc}
\usepackage[a4paper,
            inner=2.5cm,
            outer=3cm
            ]{geometry}
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off, font=small]{caption}            
\usepackage{acronym}
\usepackage{graphicx}% Include figure files
\usepackage{bm}% bold math
\usepackage{color}
\usepackage{hyperref}
\usepackage[backend=biber,
            sorting=none,
%            citestyle=numeric-comp,
%\            bibstyle=ieee, 
            style=numeric-comp,
            hyperref=true,
            maxnames=2,
            url=false,
            isbn=false,
            eprint=false,
            date=year,
            doi=false,
            giveninits=true
            ]{biblatex}
\addbibresource{references.bib}
\usepackage{xpatch}
\xpatchbibmacro{journal+issuetitle}{%
\usebibmacro{journal}%
\setunit*{\addspace}%
}{%
\usebibmacro{journal}%
\setunit*{\addcomma\space}%
}{}{}

@@

% ================================================================================================
\beginsupplement
\section{Supplementary Material}
% ================================================================================================
\subsection{General Mathematical Formalism}  \label{ssec:supplement1}
Consider the system consisting of $N$ species in $M$ different states.
In the general representation, all discussed in the main part systems  can be described with the Ordinary Differential Equations (\acp{ode}) using a non-linear function $\mbF$:
\begin{equation}
   \dot{\mbx} = \mbF(\mbx, \mbm, \mbp, \mbu),
\label{eq:model_ODE_x}
\end{equation}
where  $\mbx \in \mathbb{R}^{N}  \otimes \mathbb{R}^{M}$ is a vector describing bacterial count.
In matrix form it can be written as
\begin{equation}
    \mbx = \begin{pmatrix}
        x^1_1  & \dots & x^1_N  \\
        \vdots &       & \vdots \\
        x^M_1  & \dots & x^M_N  \\
            \end{pmatrix}
    \label{eq:model_bact}
\end{equation}
with $x_{j}^{i}$ describing the bacterial counts for species $i$ in state $j$.
$\mbu$ is the vector of different inputs of the model, e.g., temperature, pressure etc.
The vector $\mbm \in \mathbb{R}^{K}$ includes all variables corresponding to the microenvironment of the system which can be concentration of the inhibitors or activators:
\begin{equation}
    \dot{\mbm} = \mbH (\mbx, \mbm, \mbp, \mbu).
    \label{eq:model_microenv}
\end{equation}
Here function $\mbH$ is in general case also non-linear.\\
%
The advantage of the following formulation is its generality.
Depending on the knowledge about the system as well as the needed level of accuracy one can define an arbitrary number of states $M$ and control the complexity of the model.
Moreover, opposite to the already known Predictive Microbiology like Baranyi and Roberts one can easily change.\\
%
To calculate the bacterial count for one species $i$, we simply need to sum the tensor/vector $\mbx$ over all possible states:
\begin{equation}
    N_i = \sum_k \mbx_i^k.
\label{eq:N_pro_species}
\end{equation}
While the total bacterial count of all species of our system is
\begin{equation}
    N = \sum_i \sum_k \mbx_i^k.
\label{eq:N_total}
\end{equation}
%
%
%The three pool system described with equations~\ref{eq:ode_3pools} will be written as
%\begin{equation}
%    \dot \mbx  = A \mbx, \hspace{3mm} A = \begin{pmatrix}
%        -(\mu + \lambda) & 0             & 0      \\
%        \lambda          & \alpha - \mu' & 0      \\
%        \mu              & \mu'          & -\beta 
%    \end{pmatrix}.
%\label{eq:x_repres_3pools}
%\end{equation}
%
Using such matrix representation of the bacterial system, we can rewrite, for example, the three-pool model with limiting resource described with equations~\ref{eq:ode_3pool_resource} in a more compact form.
The concentrations of the pools are then the components of our variable vector $\mbx (t) = (L(t), G(t), D(t))^T$.
%
\begin{eqnarray}
    \dot \mbx  &=& \mbF(\mbx, \mbm, \mbp, \mbu) = \begin{pmatrix}
        -(\mu + \lambda) & 0               & 0      \\
        \lambda          & \alpha R - \mu' & 0      \\
        \mu              & \mu'            & -\beta 
    \end{pmatrix} \mbx\\
    \dot \mbm &=& \mbH (\mbx, \mbm, \mbp, \mbu) = -\frac{\alpha}{N_t} R G.
\label{eq:x_repres_3pools_resource}
\end{eqnarray}
Separating linear and non-linear parts in two terms yields
\begin{equation}
    \dot \mbx = \begin{pmatrix}
        -(\mu + \lambda) & 0       & 0 \\
         \lambda         & \alpha  & 0 \\
         \mu &  \mu'& - \beta 
    \end{pmatrix} 
    \mbx + \frac{\alpha}{N_t} x^2 \sum_i^3 x^i \begin{pmatrix} 0 \\ 1 \\ 0  \end{pmatrix}.
\label{eq:x_repres_3pools_resource2}
\end{equation}
Here the first term is the linear part and the second term define the quadratic behavior of the system.
%
\paragraph{Generalized Lotka-Volterra}

Likewise, the multispecies system with resource competition present described by equations~\ref{eq:model_2sp_resource_comp} can be represented in the same way by defining
\begin{equation}
    \mbm = R, \hspace{1cm}
    \mbx = \begin{pmatrix}
        L_A & L_B \\
        G_A & G_B 
    \end{pmatrix}.
\label{eq:gLV_x_m}
\end{equation}

The simpler version of one state per species (Eq.~\ref{eq:LV_simple}) can be simply written in a general mathematical form as generalized Lotka-Volterra:
\begin{equation}
    \frac{d x_i}{d t} = x_i f_i (\mbx), \hspace{3mm} \mathbf{f} = \mathbf{r} + A \mbx.
\label{eq:gLV}
\end{equation}
